caDDS: Community Assisted Distributed Database for Sequences

### 2020-06-27 Paper2020ACM/MeetingJune25.tex
June 27 5pm

Whole Paper:
1) Make it 2 column format
2) Add sir El King Morado - UP Philippine Genome Center
3) Remove references to page numbers
4) Put figures on top of a column (see what sir el king sent)

Abstract
- okay content
- fix typoL Data nods

Intro Pg 1
- Remove sentence: [the goal of this intro]
- remove subsections in inntro
- Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
- Intro should be 2 columns lang. Make each paragraph a few sentences instead.
- Intro goal: layoutht background of the work
- Flow: Idea of A lot of data -> Problems -> Proposed Solution
-

-Intro Pg 2
- 3,000,000,000 -> 3 billion
- Shorten HGP section. Combine HGP + NGS to one paragraph.
- Fix typo: Genomics is is

@@

Overall info guide: 
- tama na yung flow
- lessen info
- compress database to 1/2 sentences. Then mention centralized/distributed database.

Into: CAP
- make it 1 paragraph
- only describe the most important principle and why

Application Architecture: (Move to Theoretical Framework)
- Discuss Client Server in 1 sentence
- Discuss P2P in 1 sentence
-Describe Hybrid

-Sample: 
- First we need to discuss 2 basic architecture: CS, P2P. 
- CS is blank. 
- P2P is blank. 
- Our proposed system will be hybrid getting these aspects from CS and P2P

RRL:
- first paragraph say: We review blank. We would like to focus on blank and blank. In the end we would like to make a comparitive analysis. 
- Table Comparing the diff Databass: Remove Dat, Describe each column narratively, and describe some conclusions from the table
- Answer the question: anong halaga ng columns? 
- Discuss each column CS/P2P/H, and the data on each column
- FASTA -> Biological
- Move dat to Methodology.

Review of Related Literature:
- the review tells you you are meant to critique the work you have seen
- here summarize gaps (mej copy the problems to here)

Problem:
- summarize the gaps. To review the current implementation has problems in size.
- To summarize, centralized databases suffer from these problems. Therefore the research aims to focus on developing a database with these features
- Problem Statement: Central thesis, one unifying question. May certain level of measurability.

Objectives:
- specify metrics, activities, na once you do this succeed.
- the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
- For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed

Methodology:
- dito mapupunta yung framework and APIs
- say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.

Scope
- 2 things: 
- 1) create a theoretical design for a proposed system
- 2) evaluate the theoretical running time and speed of the operations.
- define here what you mean by proposed system and base system (FTP protocol of NCBI be specific that this is what base system means)
- base system: FTP retrieval from NCBI

Limitation
- no implement
- be transparent, time constraints and pandemic
- no security and authentication

Theoretical Framework
- discusses abstractly anong paglalaruan na ideas
- Discuss Hybrid,CS, P2P in short sentences
- Network, forego layers 1-3. Briefly discuss layers 4 and 5 and their importance.
- Discuss how retrieval operations rely on layers 4 and 5 (FTP for NCBI, torrent??)
- good flow, but too many info, need to shorten it.

Summary of the System
- call this section caDDS
- start with an outline for the entire section
- The retrieval mehanism is inspired by SeqTorr. This section describes the proposed system. This is a hybrid focused or distributed system

Community
- this section is good
- move to 7.2 or after

@@

Proof of Correctness (BIGTIME CHANGE)
- how do you guarantee that a get does what it is intended to do:
- cite preconditions, post conditions,all the paths from pre conditions to post conditions and how it would always be true
- like an if else
- shows the path from pre condition to post condition
- set pre conditions, assumptions, then show how it becomes postconditions (state after)

@@

Theoretical Analysis
- Mention base system means FTP from NCBI...
- Mention why only get/ retrieval is used. Since all operations can be broken up into small retrieval operations
- abstract size of a payload
- cite that network is assumed to be error free
- articulate assumptions better
- move the final comparisons here

Conclusion Template:
- say what you did
- summarize analysis of results
- 2 to 3 sentences of what you did/ problem
- 2 sentences correctness and speed
- correctness: system operations
- speed: faster/slower
- shown to do what is expected based on preconditions

Reccomendations
- okay

Acknowledgements -> Remove from Appendix

Biblio
- Make it APA format

>>>>

Whole Paper:
✓ 1) Make it 2 column format
✓ 2) Add sir El King Morado - UP Philippine Genome Center
✓ 3) Remove references to page numbers
✓ 4) Put figures on top of a column (see what sir el king sent)

✓ Abstract
✓ - okay content
✓ - fix typo Data nods

Intro Pg 1
✓ - Remove sentence: [the goal of this intro]
- Genomics study of blank. To generate data you need sequencing. But the data is big & needs effort. To sequence a lot, tech evolved and NGS gave high throughput
- Intro should be 2 columns lang. Make each paragraph a few sentences instead.
- Intro goal: layoutht background of the work
- Flow: Idea of A lot of data -> Problems -> Proposed Solution
- remove subsections in inntro (do this last)

-Intro Pg 2
✓ - 3,000,000,000 -> 3 billion
- Shorten HGP section. Combine HGP + NGS to one paragraph.
✓ - Fix typo: Genomics is is

@@

Overall info guide: 
✓ - tama na yung flow
- lessen info
- compress database to 1 or 2 sentences. Then mention centralized/distributed database.

Into: CAP
- shorten and make it 1 paragraph
- only describe the most important principle and why

Application Architecture: (Move to Theoretical Framework)
- Discuss Client Server in 1 sentence
- Discuss P2P in 1 sentence
- Describe Hybrid 
- First we need to discuss 2 basic architecture: CS, P2P. 
- CS is blank. 
- P2P is blank. 
- Our proposed system will be hybrid getting these aspects from CS and P2P

RRL:
- make the first paragraph say: We review blank. We would like to focus on blank and blank. In the end we would like to make a comparitive analysis. 
- Move Dat (move to theoretical framework)
- Table Comparing the diff Databass, Describe each column narratively, and describe some conclusions from the table
- Answer the question: anong halaga ng columns? 
- Discuss each column CS/P2P/H, and the data on each column
✓ - FASTA rename to Biological

Review of Related Literature:
- the review tells you you are meant to critique the work you have seen
- here summarize gaps 
✓ - copy the problems to here

80 percent ✓ Problem:
✓ - move the problems bullet points to RRL. Replace problem bullet points with a paragraph summary 
✓ - summarize the gaps. To review the current implementation has problems in size.
- To summarize, centralized databases suffer from these problems. Therefore the research aims to focus on developing a database with these features
✓ - Problem Statement: Central thesis, one unifying question. May certain level of measurability.

Objectives:
- specify metrics, activities, na once you do this succeed.
✓- the bullet points "Resolve problems blank of blank" move them to a paragraph below objectives
✓- For the paragraph below: First objective addresses these problems, second objective addresses these limitations blank.
✓- keep first 2 objectives, the last is: to evaluate the theoretical properties of the system based on the correctness and its speed
✓ - framework and APIs move to theoretical framework.

Scope
- 2 things: 
- 1) create a theoretical design for a proposed system
- 2) evaluate the theoretical running time and speed of the operations.
- define here what you mean by proposed system and base system (
- base system means FTP retrieval from NCBI

✓ Limitation
✓- make limitations section
✓- no implement
✓- be transparent, time constraints and pandemic
✓- no security and authentication

Theoretical Framework
- should have a very basic flow
- discusses abstractly anong paglalaruan na ideas
- Discuss Hybrid,CS, P2P in short sentences
- Network, forego layers 1-3. Briefly discuss layers 4 and 5 and their importance.
- Discuss how retrieval operations rely on layers 4 and 5 (FTP for NCBI, torrent??)
- good flow, but too many info, need to shorten it.
- (half check, since this still needs to be moved to theoretical framework) say we will be developing an information system. And all information systems have GET,POST,PUT,DELETE functionalities.

Summary of the System
✓ - call this section caDDS
- start with an outline for the entire section (This section will describe the different operations in ...)
- say: The retrieval mechanism is inspired by SeqTorr. This section describes the proposed system. This is a hybrid focused or distributed system

✓ Community
✓ - this section is good
✓- move to 7.2 or after

@@

Proof of Correctness (BIGTIME CHANGE + STUFF TO LEARN)
- how do you guarantee that a get does what it is intended to do:
- cite preconditions, post conditions,all the paths from pre conditions to post conditions and how it would always be true
- like an if else
- shows the path from pre condition to post condition
- set pre conditions, assumptions, then show how it becomes postconditions (state after)

@@

Theoretical Analysis
- Mention base system means FTP from NCBI (will be stated in scope but good to state again here) ...
- Mention why only get/ retrieval is used. Since all operations can be broken up into small retrieval operations
- abstract size of a payload
- cite that network is assumed to be error free
- articulate assumptions better
- move the final comparisons here

Conclusion Template:
✓ - make a conclusion section
- say what you did
- summarize analysis of results
- 2 to 3 sentences of what you did/ problem
- 2 sentences correctness and speed
- correctness: system operations
- speed: faster/slower
- shown to do what is expected based on preconditions

✓ Reccomendations
- okay

✓ Acknowledgements -> Remove from Appendix

✓ Biblio
✓ - Make it APA format

Presentation - after naaa

### 2020-06-27 Paper2020ACM/Paper2020.tex
June 27 5pm

\begin{table*}[ht]
    \caption{Comparisons of Different Databases and Technologies}
    \label{table:database_comparison_table}
    \begin{tabular}{ccccc}
    \toprule
    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ 
    \midrule
    BioTorrents & P2P           & No         & Depends                    & Biological  \\ 
    SeqTorr     & H             & No         & Fast (PH server)           & FASTA        \\ 
    PeerDB      & P2P           & No         & Depends                    & General      \\ 
    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\
    Dat Protocol    & P2P      & Yes        & Depends           & General  \\ 
    \bottomrule
    \end{tabular}
\end{table*}

@@

\section{Problem}
To restate the different gaps or problems found with other implementations of Genomic Databases
\subsection{Problems of Classical Databases}
\begin{enumerate}
\item Genome researchers download an entire database of genomics data, especially when uncertain how to update their data
\item Genome researchers only need a portion of the database, but end up downloading the entire database
\item Client-Server approach can face bottlenecks (server speed) (Section \ref{client-server})
\item P2P approach has a need of a large unmoderated active user base to work (Section \ref{p2p})

@@

\section{Objectives}

Based on the problems we found described, we propose to:

\begin{itemize}
    \item Design the framework for an information system that speeds up the download and upload of genomics data
    \begin{itemize} 
        \item To resolve 1 \& 2 of the Problems of Classical Databases. 
    \end{itemize}
    \item  Design a distributed database system
    \begin{itemize}
        \item To resolve 3 \& 4 of the Problems of Classical Databases, and 1 \& 2 of the Problems of Distributed Databases. 
    \end{itemize}
\end{itemize}

\begin{itemize}
    \item For the framework, the system should have the main API’s 
    \begin{itemize}
        \item POST genome data
        \item GET genome data (this will implement a hybrid P2P transfer)
        \begin{itemize}
            \item This will aim to speed up number 3 in the Problems of Classical Databases
        \end{itemize}
        \item PUT genome metadata (optional: PUT data)
        \item DELETE genome data
    \end{itemize}
\end{itemize}

@@

The current scope and coverage of the project will NOT consider:

\begin{itemize}
    \item Security and Authentication 
    \item Specific technologies to create the system
    \item All the data stored in the system (but focus on main data only)
    \item Implementing the actual system
\end{itemize}

@@

\subsection{Summary of the System}

@@

\subsubsection{Community Aspect} \label{community}
The system was also described as community assisted. The definition of \textbf{community} to be used here is a \textit{set of people and institutions who generally collaborate together to do research}. And the focus is the genome research community. Reemphasizing certain features from the database, that any registered user can upload data. This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community. Another important community aspect is that institutions can increase the number of data nodes. Meaning there will be an increase in the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later). Frequently used local data as well will be more easily accessible (since data nodes should have faster access speeds). The concept of closer networks having faster speeds (discussed in the network section) is at play here. Another community aspect will be this paper, publicly describing many aspects of the structure and code. So that this concept may be further developed and even implemented by genome research communities.

@@

@@

>>>>

\begin{table*}[ht]
    \caption{Comparisons of Different Databases and Technologies}
    \label{table:database_comparison_table}
    \begin{tabular}{ccccc}
    \toprule
    System      & CS/P2P/H & Available? & DL/UL speed & Type of Data \\ 
    \midrule
    BioTorrents & P2P           & No         & Depends                    & Biological  \\ 
    SeqTorr     & H             & No         & Fast (PH server)           & Biological       \\ 
    PeerDB      & P2P           & No         & Depends                    & General      \\ 
    NCBI        & CS            & Yes        & Slow (US server)           & Biological  \\
    Dat Protocol    & P2P      & Yes        & Depends           & General  \\ 
    \bottomrule
    \end{tabular}
\end{table*}

@@

To restate the different gaps or problems found with other implementations of Genomic Databases: 
\subsection{Problems of Centralized Databases}
\begin{enumerate}
\item Genome researchers download an entire database of genomics data, especially when uncertain how to update their data
\item Genome researchers only need a portion of the database, but end up downloading the entire database
\item Client-Server approach can face bottlenecks (server speed) (Section \ref{client-server})
\item P2P approach has a need of a large unmoderated active user base to work (Section \ref{p2p})

@@

\section{Problem}
To summarize the gaps described earlier, there are problems in implementations of centralized databases, specifically with bottlenecks and moderation. Other problems are specific to genome researchers with updating of the data. The last problem is the gap of usable distributed databases targeted towards genomics researchers. The main problem will be: "How to design a distributed database that is useful for genomics researchers?"

\section{Objectives}
Based on the problems we found described, we propose to:

\begin{itemize}
    \item Design the framework for an information system that speeds up the download and upload of genomics data
    \item  Design a distributed database system
    \item Evaluate the theoretical properties of the system based on the correctness and speed
\end{itemize}

The first objective addresses the genome researchers problems of centralized databases. The second objective aims to resolve the limitations of the client-server and P2P approaches, and the existing problems of distributed databases. The third objective aims to show the theoretical rationale for using these distributed databases.

@@

\section{Limitations}
The research had time constraints and physical constraints because of the unexpected Covid-19 pandemic. The research will not implement the system or explain the technology stack to be used. It will not include security and authentication.

@@

\begin{itemize}
    \item (Move to Theoretical Framework/caDDS) The system is an information system, and all information systems must have the main API’s 
    \begin{itemize}
        \item POST genome data
        \item GET genome data (this will implement a hybrid P2P transfer)
        \begin{itemize}
            \item This will aim to speed up number 3 in the Problems of Classical Databases
        \end{itemize}
        \item PUT genome metadata (optional: PUT data)
        \item DELETE genome data
    \end{itemize}
\end{itemize}

\subsection{caDDS: Community Assisted Distributed Database for Sequences}

@@

@@

\subsection{Community Aspect} \label{community}
The system was also described as community assisted. The definition of \textbf{community} to be used here is a \textit{set of people and institutions who generally collaborate together to do research}. And the focus is the genome research community. Reemphasizing certain features from the database, that any registered user can upload data. This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community. Another important community aspect is that institutions can increase the number of data nodes. Meaning there will be an increase in the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later). Frequently used local data as well will be more easily accessible (since data nodes should have faster access speeds). The concept of closer networks having faster speeds (discussed in the network section) is at play here. Another community aspect will be this paper, publicly describing many aspects of the structure and code. So that this concept may be further developed and even implemented by genome research communities.

@@

\section{Conclusion}
